---
title: "Web scraping with XPATHs"
output: html_document
---

```{r, include=FALSE}
library(rvest)
```


# Web scraping with XPATHs

Now that we've covered the low-hanging fruit ("it has an API, and a client", "it has an API") it's time to talk about what to do when a website doesn't have any access mechanisms at all - when you have to rely on web scraping. This chapter will introduce you to the rvest web-scraping package, and build on your previous knowledge of XML manipulation and XPATHs.


## Web scraping 101

__Selectors__

- Little browser extensions
- Identify the specific bit(s) you want
- Give you a unique ID to grab them with
- Not used in this course (but worth grabbing after)

__rvest__

- rvest is a dedicated web scraping package
- Makes things shockingly easy
- Read HTML page with read_html(url = \_\_\_)

__Parsing HTML__

- read_html() returns an XML document
- Use html_node() to extract contents with XPATHs

```{r, collapse=TRUE}
wiki_r <- read_html("https://en.wikipedia.org/wiki/R_(programming_language)")
wiki_r

html_node(wiki_r, xpath = "//ul")
```

## Reading HTML {.tabset .tabset-fade .tabset-pills}

### Exercise

The first step with web scraping is actually reading the HTML in. This can be done with a function from `xml2`, which is imported by `rvest` - `read_html()`. This accepts a single URL, and returns a big blob of XML that we can use further on.

We're going to experiment with that by grabbing Hadley Wickham's wikipedia page, with `rvest`, and then printing it just to see what the structure looks like.

### Instructions

- Load the `rvest` package.
- Use `read_html()` to read the URL stored at `test_url.` Store the results as `test_xml`.
- Print `test_xml`.

### script.R

```{r, collapse=TRUE}
# Load rvest
library(rvest)

# Hadley Wickham's Wikipedia page
test_url <- "https://en.wikipedia.org/wiki/Hadley_Wickham"

# Read the URL stored as "test_url" with read_html()
test_xml <- read_html(test_url)

# Print test_xml
test_xml
```

## Extracting nodes by XPATH {.tabset .tabset-fade .tabset-pills}

### Exercise

Now you've got a HTML page read into R. Great! But how do you get individual, identifiable pieces of it?

The answer is to use `html_node()`, which extracts individual chunks of HTML from a HTML document. There are a couple of ways of identifying and filtering nodes, and for now we're going to use XPATHs: unique identifiers for individual pieces of a HTML document.

These can be retrieved using a browser gadget we'll talk about later - in the meanwhile the XPATH for the information box in the page you just downloaded is stored as `test_node_xpath`. We're going to retrieve the box from the HTML doc with `html_node()`, using `test_node_xpath` as the `xpath` argument.

### Instructions

- Use `html_node()` to retrieve the node with the XPATH stored at `test_node_xpath` from `test_xml` document you grabbed in the last exercise.
- Print the first element of the results.

### script.R

```{r, collapse=TRUE}
test_node_xpath <- "//*[contains(concat( \" \", @class, \" \" ), concat( \" \", \"vcard\", \" \" ))]"

# Use html_node() to grab the node with the XPATH stored as `test_node_xpath`
node <- html_node(x = test_xml, xpath = test_node_xpath)

# Print the first element of the result
node[1]
```

## HTML structure

__Tags__

- HTML is content within tags 
- Like XML
- <p> this is a test </p>

__Attributes__

`<a href = "https://en.wikipedia.org/"> this is a test </a>`

__Extracting information__

- html_text(x = \_\_\_) \- get text contents
- html_attr(x = \_\_\_, name = \_\_\_) \- get specific attribute 
- html_name(x = \_\_\_) \- get tag name

## Extracting names {.tabset .tabset-fade .tabset-pills}

### Exercise

The first thing we'll grab is a name, from the first element of the previously extracted table (now stored as `table_element`). We can do this with `html_name()`. As you may recall from when you printed it, the element has the tag `<table>...</table>`, so we'd expect the name to be, well, `table`.

### Instructions

- Extract the name of `table_element` using the function `html_name()`. Save it as `element_name`.
- Print `element_name`.

### script.R

```{r, collapse=TRUE}
# Use html_node() to grab the node with the XPATH stored as "table_element"
table_element <- html_node(test_xml, xpath = "//table")

# Extract the name of table_element
element_name <- html_name(table_element)

# Print the name
element_name
```

## Extracting values {.tabset .tabset-fade .tabset-pills}

### Exercise

Just knowing the type of HTML object a node is isn't much use, though (although it can be very helpful). What we really want is to extract the actual text stored within the value.

We can do that with (shocker) `html_text()`, another convenient `rvest` function that accepts a node and passes back the text inside it. For this we'll want a node within the extracted element - specifically, the one containing the page title. The xpath value for that node is stored as `second_xpath_val`.

Using this xpath value, extract the node within `table_element` that we want, and then use `html_text` to extract the text, before printing it.

### Instructions

- Extract the element of `table_element` referred to by `second_xpath_val` and store it as `page_name`.
- Extract the text from `page_name` using `html_text()`, saving it as `page_title`.
- Print `page_title`.

### script.R

```{r, collapse=TRUE}
second_xpath_val <- "//*[contains(concat( \" \", @class, \" \" ), concat( \" \", \"fn\", \" \" ))]"

# Extract the element of table_element referred to by second_xpath_val and store it as page_name
page_name <- html_node(x = table_element, xpath = second_xpath_val)

# Extract the text from page_name
page_title <- html_text(page_name)

# Print page_title
page_title
```

## Test: HTML reading and extraction {.tabset .tabset-fade .tabset-pills}

### Question

Time for a quick test of what we've learned about HTML. What would you use to extract the type of HTML tag a value is wrapped in?

### Answer

Possible Answers

[x] html_attr()      
[o] html_name()

## Reformatting Data 

HTML tables

- HTML tables are dedicated structures: <table>...</table>
- They can be turned into data.frames with html_table()
- Use colnames(table) <- c("name", "second_name") to name the columns

Turning things into data.frames

- Non-tables can also become data.frames
- Use data.frame(), with the vectors of text or names or attributes

## Extracting tables {.tabset .tabset-fade .tabset-pills}

### Exercise

The data from Wikipedia that we've been playing around with can be extracted bit by bit and cleaned up manually, but since it's a table, we have an easier way of turning it into an R object. `rvest` contains the function `html_table()` which, as the name suggests, extracts tables. It accepts a node containing a table object, and outputs a data frame.

Let's use it now: take the table we've extracted, and turn it into a data frame.

### Instructions

- Turn `table_element` into a data frame and assign it to `wiki_table`.
- Print the resulting object.

### script.R

```{r, collapse=TRUE}
# Turn table_element into a data frame and assign it to wiki_table
wiki_table <- html_table(table_element)

# Print wiki_table
wiki_table
```

## Cleaning a data frame {.tabset .tabset-fade .tabset-pills}

### Exercise

In the last exercise, we looked at extracting tables with `html_table()`. The resulting data frame was pretty clean, but had two problems - first, the column names weren't descriptive, and second, there was an empty row.

In this exercise we're going to look at fixing both of those problems. First, column names. Column names can be cleaned up with the `colnames()` function. You call it on the object you want to rename, and then assign _to that call_ a vector of new names.

The missing row, meanwhile, can be removed with the `subset()` function. `subset` takes an object, and a condition. For example, if you have a data frame df containing a column x, you could run

`subset(df, !x == "")`

to remove all rows from `df` consisting of empty strings (`""`) in the column `x`.

### Instructions

- Rename the columns of `wiki_table` to `"key"` and `"value"` using `colnames()`.
- Remove the empty row from `wiki_table` using `subset()`, and assign the result to `cleaned_table`.
- Print `cleaned_table`.

### script.R

```{r, collapse=TRUE}
# Rename the columns of wiki_table
colnames(wiki_table) <- c("key", "value")

# Remove the empty row from wiki_table
cleaned_table <- subset(wiki_table, !value == "")

# Print cleaned_table
cleaned_table
```
